%!TEX root = ../template.tex
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% acknowledgements.tex
%% NOVA thesis document file
%%
%% Text with acknowledgements
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\typeout{NT FILE acknowledgements.tex}%

\begin{ntacknowledgements}

This work is far from being only mine. It is the combined effort of all people and entities that followed me throughout this journey. Nothing here would have been possible without them. With all of you I shared doubts, frustrations, but also many victories. To all of you, I give my sincere aknowledgement.

I would like to start by thanking the \textit{Faculty of Science and Technology from the NOVA University of Lisbon}, for giving me the environment and conditions to pursue this path, as well as to the financing entities, \textit{Fundação para a Ciência e Tecnologias} and \textit{Volkswagen Autoeuropa} for supporting this work. I also want to thank the Hanse-Wissenschaftskolleg for supporting my stay at Bremen and providing me an incredible experience and the Fullbright Comission for making my stay in Riverside possible.

A special thank you has to go to my family from the Biosignals Group. In this laboratory, I spent many challenging moments, but it was always a safe place where I could count on any of my dear fellows. I would start to acknowledge my mentor, prof. Hugo Gamboa, the main reason why I continued working in science. Although filled with doubts, prof. Hugo convinced me this was the right path. He was and always will be a reference, whom valued, inspired and supported me, but above all, challenged me in doing better and achieving more. Of the many challenges, some were not fulfilled and stayed as abstract ideas. But many others matured and bloomed into concrete works presented here in this manuscript. Prof. Hugo helped me grow as a researcher but also helped me shape part of who I am today. For all this, I am very grateful, hoping he enjoyed the ride as much as I did. David became a friend and older \textit{brother}, showing support and companionship. He was essential in my early career. Thank you for all the advises, good conversations, and cheerful moments we spent together. Who thought that all would start by painting the lab furniture...To Cátia, central figure of the group, we shared so many good moments since the creation of the lab with so much \textit{bricolage}. She taught me to make things easier and make better decisions, also being an important confident. Daniel, a source of knowledge, but mostly, a source of skepticism, highly necessary to challenge and test us in improving our work. I fondly remember our great trip to Denmark, thank you for everything. Ricardo, thank you for the energy and motivation that you shared so nicely everyday you went to work. Nafiseh, thank you to make me know myself better. Phillip, thank you for all the wonderful conversations and discussions about so many topics (work and non-work related). Luís, such a motivating and energetic figure. Thank you for sharing so much knowledge, listening to my outbursts and giving me so many rides to Porto Brandão. To the master students I had the luck to follow or collaborate with, Guilherme, Diogo, António, Sara Santos and Sara Silva, Lua and Rita, thank you for teaching me so much. All the remaining members, Mariana and Daniel, I hope you can enjoy your time in the group and can continue the good work being done. I also want to thank our \textit{old} neighbours, Paulo and Jorge. Thank you for your companionship, good mood and constant motivation. You were important pieces throughout these years.

I had the luck to share this journey with many, having collaborate with multiple entities. A special thank you goes to Fraunhofer-AICOS, specially to Duarte, for all the knowledge shared, your rigor, excellence, and work culture. Thank you also for all our enjoyable conversations about time series. Thank you to PLUX for all the equipment shared. Autoeuropa Volkswagen, thank you to the Industrial and Management Department, namely Pedro, Zé, Natércia, Jakie and specially Carlos, thank you for your availability and desire to make your ideas work. A big thank you goes to the Time Series group from the University of California, Riverside. Ryan, Audrey, Renjie and Maryam, as well as to Rutuja and Jiasi, thank you so much for the great environment you provided, my stay would not have been so pleasant without you. A special thank you goes to Eamonn Keogh. You are a great scientific inspiration. I always followed your work with a great interest and admiration. Thank you for receiving me so well, sharing so much knowledge and providing me such an amazing experience. I also had the chance to stay at the Cognitive Systems Lab from the Universität Bremen. It was a great pleasure and honour to work with such a wonderful group. Thank you to all the team members for receiving and treating me so well. A special thank you to Prof. Tanja, Felix, Rafael, Moritz, Yale and Hui. I will never forget the ideas, knowledge and great activities we shared. It was such a wonderful experience being at the CSL and I will forever cherish it. Thank you!

I kept the most dear to me for last. Many of them had to endure bits of my frustrations, but all helped to relieve it when gathering and enjoying our time together. Francisco, Nuno and Gonçalo, thank you for always being there when needed. My dear parents and sister, José, Angela and Nádia, you always guided me in the right way. I am sorry for being sometimes distant, and can't thank you enough for everything you did and always do for me. Ana, my dear wife, you lived and endured this as much as I did. I have to apologize for all the time I was \textit{too busy} with something else. I can't thank you enough for your patience, moral support, and guidance, for joining me in this journey, and for simply being there when I needed the most, making everything easier. My achievements are yours too.

\end{ntacknowledgements}