%!TEX root = ../template.tex
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% chapter2.tex
%% NOVA thesis document file
%%
%% Chapter with the template manual
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\typeout{NT FILE chapter2.tex}%

\chapter{Sensing the Physical World}
\label{cha:theory}

\section{Sensing the Physical World}
\label{sec:sensing_physical}

- Bring an overview of the sensors used for a general set of things - motion, physiological, etc...

- These measure the physical world and retrieve physical changes in what they measure

- These changes can be related with something meaningful and relevant that occurred and can be seen on the data

- In this work, we apply the methods in all kinds of scenarios to show their agnosticism to domains, but focus our attention to a specific context where introducing sensing technology can have a strong impact.



\section{Occupational Variables in the Industry}
\label{subsec:occup_var}


- Occupational variables that affect worker's health have long been studied and already defined in several screening tools from standard ergonomic guidelines. We can name EAWS, OCRA, RULA, etc...

- These worksheets are a reference for ergonomists in identifying the variables of interest to measure the risk of each activity performed by a worker. 

- The multiple set of actions of a workstation can be analyzed

- Typically, these variables are related with motion and posture of body segments. Several scenarios are studied and variables extracted are frequency, intensity and duration of activity. Study specific activities, measure the risk based on vibration from machine or tools (https://www.cdc.gov/niosh/topics/ergonomics/ergoprimer/default.html)

- All these variables, being related with motion, should be studied  

definition 1 - workstation

definition 3 - intensity, duration and frequency

definition 2 - vibration

- hand - is a special case. We should measure the vibration on the grip 


\subsection{Sensing Worker's Health}
\label{subsec:sensing_worker}

The type of variables that are required to perform a risk assessment are related with motion, posture and vibration. These are physical variables that can be quantified by means of inertial sensors, such as accelerometer, gyroscope and magnetometer. These three sensors are used together to compensate limitations of each other in error accumulation from sensor drifting.

- With these sensors, we can measure the orientation of body segments in terns of other body segments or standard body planes (sagital or frontal plane). 